Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-03-26 arcticgrids.tex
new elev figure
melt bias plot

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change. Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a). Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain. With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past. In this study a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), including brand new variable-resolution grids, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the climate and surface mass balance of the Greenland Ice Sheet.

In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs. These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g. horizontal derivatives) in the equation set can be solved for exactly. While spectral transform methods are still used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}. Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions. This additional damping reduces the effective resolution in polar regions, and the resolved scales are approximately the same everywhere on the grid.

@@

The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}. For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale Polar Lows are not well resolved at these resolutions. These mesoscale systems are prevalent during the cold season and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface. The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}). While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions. For example, GrIS precipitation is typically confined to the ice-sheet margins (predominately the southeastern margin) where orographic precipitation is generated by steep topographic slopes. GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation. GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.

The goal of this study is to characterize the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in different ways. The manuscript is laid out as follows: Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study. The Arctic refined grids were developed by the authors, and this section serves as their official documentation in CESM2.2. Section~\ref{sec:methods} also contains a description of the experiments along with reanalysis datasets and post-processing software used for evaluating the model simulations. Section~\ref{sec:results} contains the results of the experiments, followed by Section~\ref{sec:conclusions} that provides a general discussion and conclusions.

@@

Six grid are evaluated in this study (Table~\ref{tbl:table1}). The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b). The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical paramerizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.

Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic (Figure~\ref{fig:vr-grids}). The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $ARCTIC$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region. The $ARCTICGRIS$ grid is identical to the $ARCTIC$ grid, but contains an additional patch covering the big island of Greenland with $\frac{1}{8}^{\circ}$ resolution.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccccc}
   \hline
   grid name & dycore & $\Delta x_{eq}$ (km) & $\Delta x_{refine}$ (km) & $\Delta t_{phys}$ (s) \\ 
   \hline
   $f19$ & FV & 278 & - &1800 \\
   $f09$ & FV & 139 & - &1800 \\
   $ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
   $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
   $ARCTIC$ & SE & 111 & 28 & 450 \\
   $ARCTCIGRIS$ & SE & 111 & 14 & 225 \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{eq}$ refers to average equatorial grid spacing, $\Delta x_{refine}$ refers to grid spacing in the refined region (if applicable) and $\Delta t_{phys}$ refers to the physics time-step. The dycore abbreviation FV refers to the finite-volume dycore, SE the spectral-element dycore and SE-CSLAM the spectral-element dycore w/ CSLAM tracer advection.}
 \label{tbl:table1}
 \end{table*}

@@

The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used in all simulations in this study. CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme. CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions. Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} inlcuding convective momentum transport \cite{RSG2010JAS}. PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.

All grids and dycores in this study use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$. The physics time-step is dependent on grid resolution. Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step is reduced to avoid large time truncation errors \cite{HR2018JAMES}. The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids (Table~\ref{tbl:table1}.

@@

Several observational datasets are used in this study to understand the performance of the simulations. A list of the datasets used in this study are shown in Table~\ref{tbl:table2}. Several of these products (ERA5, CALIPSO and CERES) are near-global gridded datasets commonly used to evaluate GCMs. Surface mass balance datasets are gathered from several sources. RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA interim and ERA5 renalysis at its domain boundaries. The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}. The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit} maintains a repository of snow pit and and ice core SMB measurements, as well as the IceBridge radar accumulation dataset. The LIVVkit dataset is compared against model simulations by finding the nearest grid cell center to the location of each observation.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccc}
   \hline
   data product & years used in this study & citation \\ 
   \hline
   ERA5 & 1979-1998 & \cite{ERA5} \\
   CERES-EBAF ED4.1 & 2003-2020 & \cite{CERES-EBAF} \\
   CALIPSO-GOCCP & 2006-2017 & \cite{CALIPSO-GOCCP} \\
   RACMO2.3 & 1979-1998 & \cite{NETAL2015TC} \\
   RACMO2.3p2 & 1979-1998 & \cite{NETAL2019SCIENCE} \\
   LIVVkit,v2 & 1979-1998 & \cite{LIVVkit} \\
 \hline
 \end{tabular}
  \caption{Description of observational datasets used in this study.}
 \label{tbl:table2}
 \end{table*}

@@

A common high resolution dataset is used to generate the GrIS boundary conditions in all grids, using the CLM dataset creation tools. Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS. Figure~\ref{fig:grisdx} shows the GrIS ice mask area across the different grids, as a function of the number of grid points. Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset. RACMO2.3, however, uses a smaller ice mask, about $3\%$ smaller than the raw ice mask dataset.

Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, but we seek to compare these integrated quantities to RACMO2.3. Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks. Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$, the lowest resolution grid for both dycores considered in this study. In addition, two remapping algorithms are used; ESMF conservative and TempestRemap high-order, monotone algorithm. In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_grisres.pdf}
\end{center}
\caption{The spatial properties of the GrIS as represented by different grids in this study. (Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks.}
\label{fig:grisdx}
\end{figure}

@@

Before delving into the simulated characteristics of the Arctic, the global mean differences between the various grids and dycores are assessed. Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between the uniform resolution grids and dycores. The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere. This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB. The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the low and mid-latitudes in the $f09$ grid.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-lores.pdf}
\end{center}
\caption{.}
\label{fig:dT-lores}
\end{figure}

As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}). The stratosphere has a different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes. Figure~\ref{fig:dT-lores} also shows differences in temperature between $ne30pg3$ and $ne30pg2$, which are small, although there is a slight warming near the tropopause at high latitudes. This is consistent with the similar climates found between these grids in \cite{HETAL2019JAMES}.

Comparing the variable-resolution grids to the uniform resolution grids is complicated because we simultaneously increase the grid resolution and reduce the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}. An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$. Figure~\ref{fig:dT-hires} shows the change in the climatological summer temperatures in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$. A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB. Figure~\ref{fig:dT-hires} shows the difference in climatological summer temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid. The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-hires.pdf}
\end{center}
\caption{.}
\label{fig:dT-hires}
\end{figure}

It's useful to understand summer temperature biases, instead of annual means, due to its control on ice/snow melt \cite{O2001JAM,HT2008Paleo}. Figure~\ref{fig:dThyps} shows the 1979-1998 lower troposphere summer temperature bias relative to ERA5. It is computed from the 500 hPa-1000 hPa geopotential thickness, solving for the layer mean virtual temperature using the hypsometric equation. The results generally track with the analysis of the zonal mean height plots; increasing resolution from $f19$ to $f09$ leads to a warmer climate, and the $1^{\circ}$ spectral-elements grids are warmer than the finite-volume grids. The summer temperatures in the finite-volume grids are persistently colder than ERA5 at high latitudes, whereas the $1^{\circ}$ spectral-element grids are warmer than ERA5 at only very high-latitudes, north of $80^{\circ}$. All grid illustrate a north-south gradient in bias over Greenland, in which the summer temperature bias becomes more positive in the northward direction. This pattern is also evident in the 2m summer temperature bias over Greenland (not shown).

The $ARCTIC$ grid has similar summer temperatures to the $1^{\circ}$ spectral-element grids, but it is a bit warmer over northern Eurasia and the North Pole. An anomalous cooling patch forms to the west of Greenland, centered over Baffin Island. The $ARCTICGRIS$ grid is warmer than the $ARCTIC$ grid over most of the Arctic region, but approximately maintains the same pattern of summer temperature bias as in the $ARCTIC$ grid.

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffERA5_Thyps.pdf}
\end{center}
\caption{.}
\label{fig:dThyps}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffCERES_SWCF.pdf}
\end{center}
\caption{.}
\label{fig:SWCF}
\end{figure}

@@

%\begin{figure}[t]
%\begin{center}
%         \includegraphics[width=130mm]{figs/temp_transect_obsperiod.pdf}
%\end{center}
%\caption{.}
%\label{fig:trans}
%\end{figure}

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=130mm]{figs/temp_ktransect.pdf} \\
         \includegraphics[width=130mm]{figs/temp_btransect.pdf} \\
\end{tabular}
\end{center}
\caption{.}
\label{fig:ztrans}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/Vis1923.jpg}
\end{center}
\caption{.}
\label{fig:viz}
\end{figure}

>>>>

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change. Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a). Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain. With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past. In this study a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including brand new variable-resolution grids, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the climate and surface mass balance of the Greenland Ice Sheet.

In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs. These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g. horizontal derivatives) in the equation set can be solved for exactly. While spectral transform methods are still used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems. The pole problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}. Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions. This additional damping reduces the effective resolution in polar regions such that the resolved scales are approximately the same everywhere on the grid.

@@

The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}. For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale Polar Lows are not well resolved at these resolutions. These mesoscale systems are prevalent during the cold season and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface. The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}). While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions. For example, GrIS precipitation is concentrated at the ice-sheet margins, where orographic precipitation is generated by steep topographic slopes. GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation. GCMs struggle to resolve the magnitude and extent of these features \cite{P2010CC,VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.

The goal of this study is to characterize the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in different ways. The manuscript is laid out as follows: Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study. The Arctic refined grids were developed by the authors, and this section serves as their official documentation in CESM2.2. Section~\ref{sec:methods} also contains a description of the experiments along with the datasets and methods for evaluating the model simulations. Section~\ref{sec:results} contains the results of the experiments, followed by Section~\ref{sec:conclusions} that provides a general discussion and conclusions.

@@

Six grid are evaluated in this study (Table~\ref{tbl:table1}). The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b). The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics. An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical parameterizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.

Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic (Figure~\ref{fig:vr-grids}). This manuscript serves as the official documentation of these grids. The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $ARCTIC$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region. The $ARCTICGRIS$ grid is identical to the $ARCTIC$ grid, but contains an additional patch covering the big island of Greenland with $\frac{1}{8}^{\circ}$ resolution.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lccccc}
   \hline
   grid name & dycore & $\Delta x_{eq}$ (km) & $\Delta x_{refine}$ (km) & $\Delta t_{phys}$ (s) \\ 
   \hline
   $f19$ & FV & 278 & - &1800 \\
   $f09$ & FV & 139 & - &1800 \\
   $ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
   $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
   $ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 \\
   $ARCTIC$ & SE & 111 & 28 & 450 \\
   $ARCTCIGRIS$ & SE & 111 & 14 & 225 \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{eq}$ refers to average equatorial grid spacing, $\Delta x_{refine}$ refers to grid spacing in the refined region (if applicable) and $\Delta t_{phys}$ refers to the physics time-step. The dycore abbreviation FV refers to the finite-volume dycore, SE the spectral-element dycore and SE-CSLAM the spectral-element dycore w/ CSLAM tracer advection.}
 \label{tbl:table1}
 \end{table*}

The physics time-step should depend on grid resolution. Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step is reduced to avoid large time truncation errors \cite{HR2018JAMES}. The $ARCTIC$ and $ARCTICGRIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in the standard $1^{\circ}-2^{\circ}$ grids (Table~\ref{tbl:table1}).

All grids and dycores in this study use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$. However, any grid or dycore can in principle be run with a higher-top and/or with finer vertical resolution.

@@

The CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{CAM63} is used in all simulations in this study. CAM version 6 physics and later is most notably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme. CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions. Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}. PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.

@@

Several observational datasets are used in this study to understand the performance of the simulations. A list of the datasets used in this study are shown in Table~\ref{tbl:table2}. Several of these products (ERA5, CALIPSO and CERES) are near-global gridded datasets commonly used to evaluate GCMs. Surface mass balance datasets were initially gathered from several sources. RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA interim and ERA5 renalyses at its domain boundaries. The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.

In-situ SMB (snow pit and ice cores) and radar accumulation datasets (e.g., IceBridge) are conveniently maintained in The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit}. However, these point-wise measurements are difficult to compare to model output spanning several different grids, and would be more useful if the surface mass balance from the individual EC's were available from the model output, which they are not. A nearest neighbor technique was used to do an initial analysis, and indicated that the model biases are similar to those computed using the RACMO datasets. Owing to the uncertainty of comparing gridded fields to point-wise measurements, and the lack of additional information it adds with regard to model biases, these datasets were omitted from the analysis in this study.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccc}
   \hline
   data product & years used in this study & resolution & citation \\ 
   \hline
   ERA5 & 1979-1998 & $\frac{1}{4}^{\circ}$ & \citeA{ERA5} \\
   CERES-EBAF ED4.1 & 2003-2020 & $1^{\circ}$ & \citeA{CERES-EBAF} \\
   CALIPSO-GOCCP & 2006-2017 & $1^{\circ}$ & \citeA{CALIPSO-GOCCP} \\
   RACMO2.3 & 1979-1998 & 11 km & \citeA{NETAL2015TC} \\
   RACMO2.3p2 & 1979-1998 & 5.5 km & \citeA{NETAL2019SCIENCE} \\
 \hline
 \end{tabular}
  \caption{Description of observational datasets used in this study.}
 \label{tbl:table2}
 \end{table*}

@@

A common high resolution dataset is used to generate the GrIS boundary conditions in all grids, using the CLM dataset creation tools. Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS. Figure~\ref{fig:grisdx} shows the GrIS ice mask area errors (computed as percent differences from the high resolution dataset) across the different grids, and as a function of the number of grid points. Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset. RACMO2.3, however, uses a smaller ice mask, about $3\%$ smaller than the raw ice mask dataset.

The area errors in Figure~\ref{fig:grisdx} don't seem large, but even small changes in ice mask area can lead to large differences in integrated SMB \cite{HETAL2022TC}. Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks. Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ grids, the lowest resolution grid for each dycore considered in this study. In addition, two remapping algorithms are used; Earth System Modeling Framework \cite<ESMF;>{ESMF} first-order conservative and TempestRemap \cite{TempestRemap} high-order, monotone algorithm. In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_grisres.pdf}
\end{center}
\caption{The spatial properties of the GrIS as represented by different grids in this study. (Left) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
\label{fig:grisdx}
\end{figure}

@@

Before delving into the simulated characteristics of the Arctic, the global mean differences between the various grids and dycores are assessed. Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between the uniform resolution grids and dycores. The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere. This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \citeA{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB. The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the low and mid-latitudes in the $f09$ grid.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-lores.pdf}
\end{center}
\caption{1979-1998 annual mean (left column) temperature and (right column) CLUBB temperature tendencies in zonal mean height space, expressed as differences between the various $1^{\circ}-1^{\circ}$ grids. The thick grey and magenta lines are the tropopause for control run and the test run, respectively.}
\label{fig:dT-lores}
\end{figure}

As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}). The stratosphere has a uniquely different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes. Figure~\ref{fig:dT-lores} also shows differences in temperature between $ne30pg3$ and $ne30pg2$, which are small, although there is a slight warming near the tropopause at high latitudes. This is consistent with the similar climates found between these grids in \citeA{HETAL2019JAMES}.

Comparing the variable-resolution grids to the uniform resolution grids is complicated because we simultaneously increase the grid resolution and reduce the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}. An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$. Figure~\ref{fig:dT-hires} shows the change in the climatological summer temperatures in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$. A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB. Figure~\ref{fig:dT-hires} shows the difference in climatological summer temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid. The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the comparison.

Figure~\ref{fig:dT-hires} indicates that the $ARCTICGRIS$ grid is a lot warmer than the $ARCTIC$ grid in the Arctic. Part of this may be due to the shorter physics time-step; a similar physics-time-step sensitivity in the $ne30pg3^{*}$ comparison is evident in the CLUBB tendencies in the $ARCTCGRIS$ grid. However, the temperature response is disproportionately large relative to the change in CLUBB tendencies. This summer warming appears to be a result of variations in the stationary wave pattern, with anomalous southerly winds occurring to the west of Greenland (not shown). This dynamic response is extremely interesting because other than the physics time-step, the only difference between $ARCTICGRIS$ and the $ARCTIC$ grid is the doubling of resolution of Greenland. This behavior will be explored more fully in a future study.

\begin{figure}[t]
\begin{center}
         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-hires.pdf}
\end{center}
\caption{As in Figure~\ref{fig:dT-lores} but for the short-time step experiment and the variable-resolution grids. The fields plotted are for the climatological northern hemisphere summer. The focus on summer is because that is when the resolution response is largest, and the refined regions are located in the northern hemisphere.}
\label{fig:dT-hires}
\end{figure}

It's useful to understand summer temperature biases due to its control on ice/snow melt over the Greenland Ice Sheet \cite<GrIS;>{O2001JAM}. Figure~\ref{fig:dThyps} shows the 1979-1998 lower troposphere summer temperature bias relative to ERA5. It is computed from the 500 hPa-1000 hPa geopotential thickness, solving for the layer mean virtual temperature using the hypsometric equation. The results generally track with the analysis of the zonal mean height plots; increasing resolution from $f19$ to $f09$ leads to a warmer climate, and the $1^{\circ}$ spectral-elements grids are warmer than the finite-volume grids. The summer temperatures in the finite-volume grids are persistently colder than ERA5, whereas the $1^{\circ}$ spectral-element grids are not as cold, and are actually warmer than ERA5 at very high-latitudes, north of $75^{\circ}$. All grids illustrate a north-south gradient in bias over Greenland, in which the summer temperature bias becomes more positive on the northern side of the ice sheet. This pattern is also evident in the 2m summer temperature bias over Greenland (not shown).

The $ARCTIC$ grid has similar summer temperatures to the $1^{\circ}$ spectral-element grids, but it is a bit warmer over northern Eurasia and the North Pole. An anomalous cooling patch forms to the west of Greenland, centered over Baffin Island. The $ARCTICGRIS$ grid is warmer than the $ARCTIC$ grid over most of the Arctic, but maintains a similar spatail pattern of summer temperature bias as in the $ARCTIC$ grid.

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffERA5_Thyps.pdf}
\end{center}
\caption{1979-1998 lower troposphere, northern hemisphere summer virtual temperature biases, computed as the difference from ERA5. Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation. ERA5 is very high-resolution at high-latitudes owing to the use of a $\frac{1}{4}^{\circ}$ latitude-longitude grid, and so all model differences are computed after mapping ERA5 to their respective model grids.}
\label{fig:dThyps}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{figs/temp_contours_diffCERES_SWCF.pdf}
\end{center}
\caption{1979-1998 northern hemisphere summer shortwave cloud forcing bias, relative to the CERES-EBAF gridded dataset. Differences are computed after mapping all model output to the $1^{\circ}$ CERES-EBAF grid.}
\label{fig:SWCF}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=130mm]{figs/temp_zsects.pdf}
\end{tabular}
\end{center}
\caption{.}
\label{fig:ztrans}
\end{figure}

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=100mm]{figs/temp_xy_diffRACMO_melt.pdf}
\end{tabular}
\end{center}
\caption{.}
\label{fig:bias}
\end{figure}

@@

%\begin{figure}[t]
%\begin{center}
%         \includegraphics[width=130mm]{figs/Vis1923.jpg}
%\end{center}
%\caption{.}
%\label{fig:viz}
%\end{figure}
